---
title: "Class notes"
author: "Risa Dickerman"
date: "2026-10-01"
output: github_document
---


```{r}
library(tidyverse)
library(ggridges)

```

```{r}
library(p8105.datasets)
data("weather_df")

```
Now we have everything we need!

##Scatter plots!

Create my first scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()

```
geom_point() = scatterplot

I always like the dataframe first
```{r}
ggp_temp_scatterplot =
weather_df |> 
  ggplot(aes(x = tmin, y =tmax)) +
  geom_point()
```
If you want to save plots name it and export it in a different way. One base plot and save it in a different way
Same as 40.

Make it prettier
```{r}
ggp_temp_scatterplot =
weather_df |> 
  ggplot(aes(x = tmin, y =tmax, color = name)) +
  geom_point()
ggp_temp_scatterplot
```
alpha = alpha blending. Just makes points more transparent.
Add geomsmoothe to create a line through the dataset

```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE)
weather_plot
```



```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(aes(color = name), alpha = 0.25) + 
  geom_smooth(se = FALSE)
weather_plot
```
where you put aesthetics matters. 
Aesthetics are up to you

```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
weather_plot
```
Don't have to have a scatterplot for example above. Can just have the mean lines

Show faceting

```{r}
gg_weather = 
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(. ~ name)
gg_weather
```
Splitting plots on name from dataset.

Both above and below work the same
```{r}
gg_weather = 
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid( cols = vars(name))
gg_weather
```

Let's look at something else

```{r}
weather_df |> 
   ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```

make a plot of central park tmax v tmin only, and convert temperatures to fahrnenheit.
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(tmax = 9/5 *tmax +32,
    tmin = 9/5 *tmin +32) |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

What's a hex plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

## Univariate plots
Histogram
```{r}
weather_df |> 
  ggplot(aes(x = tmax)) +
  geom_histogram()
```

Create histogram with different names
```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram()+
  facet_grid(. ~ name)
```
color around edges bc for histogram the color is the edge of the bar. instead do fill to get whole one
position(dodge)
Just visual comparison group to group. Reading horizontally is harder than reading otherways. So not the best for facet_grid

#Density plots
Density plots are great!! (Density = smoothed out histogram)

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
    geom_density(alpha = 0.3)
```
Easier to see what the overlapping peices are. See each one individually and as a group
Can do color outside or fill but it will cover up. Use alpha to make things transparent

#Boxplots

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

#Violin plot

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```
Violin shows you a little bit more information than boxplots. Sometimes bimodal distribution but won't be able to pick up on it


#Ridge plots
```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```
Ridge plots - imagine starting with a density, what's showing up here is just the overall density of everything.
Could stack  or could make the much more exciting ridges. Once overlapping with color. With many data hard to look at everything so it's more useful to separate them out rather than just density plots
