---
title: "Class notes"
author: "Risa Dickerman"
date: "2026-10-01"
output: html_document
---


```{r}
library(tidyverse)
library(ggridges)

```

```{r}
library(p8105.datasets)
data("weather_df")

```
Now we have everything we need!

##Scatter plots!

Create my first scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()

```
geom_point() = scatterplot

I always like the dataframe first
```{r}
ggp_temp_scatterplot =
weather_df |> 
  ggplot(aes(x = tmin, y =tmax)) +
  geom_point()
```
If you want to save plots name it and export it in a different way. One base plot and save it in a different way
Same as 40.

Make it prettier
```{r}
ggp_temp_scatterplot =
weather_df |> 
  ggplot(aes(x = tmin, y =tmax, color = name)) +
  geom_point()
ggp_temp_scatterplot
```
alpha = alpha blending. Just makes points more transparent.
Add geomsmoothe to create a line through the dataset

```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE)
weather_plot
```



```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(aes(color = name), alpha = 0.25) + 
  geom_smooth(se = FALSE)
weather_plot
```
where you put aesthetics matters. 
Aesthetics are up to you

```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
weather_plot
```
Don't have to have a scatterplot for example above. Can just have the mean lines

Show faceting

```{r}
gg_weather = 
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(. ~ name)
gg_weather
```
Splitting plots on name from dataset.

Both above and below work the same
```{r}
gg_weather = 
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid( cols = vars(name))
gg_weather
```

Let's look at something else

```{r}
weather_df |> 
   ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

```

make a plot of central park tmax v tmin only, and convert temperatures to fahrnenheit.
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(tmax = 9/5 *tmax +32,
    tmin = 9/5 *tmin +32) |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```


