---
title: "01_vizualization"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)

```

```{r}
library(p8105.datasets)
data("weather_df")

```
Now we have everything we need!

##Scatter plots!

Create my first scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()

```
geom_point() = scatterplot

I always like the dataframe first
```{r}
ggp_temp_scatterplot =
weather_df |> 
  ggplot(aes(x = tmin, y =tmax)) +
  geom_point()
```
If you want to save plots name it and export it in a different way. One base plot and save it in a different way
Same as 40.
New approach, same plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax))+
  geom_point()
```

Make it prettier
```{r}
ggp_temp_scatterplot =
weather_df |> 
  ggplot(aes(x = tmin, y =tmax, color = name)) +
  geom_point()
ggp_temp_scatterplot
```
alpha = alpha blending. Just makes points more transparent.
Add geomsmoothe to create a line through the dataset

See and edit a plot object

```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot....
Start with the same one and make it fancy
```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE)
weather_plot
```

What about the 'aes' placement ?
Although ggplot knows the axes, but it knows that aesthetics is defined in the points.
Aesthetics defined can be defined here or can change placement
```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) + 
  geom_smooth()
weather_plot
```

Let's facet some things:

```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.2) + 
  geom_smooth(se = FALSE, size = 1) +
  facet_grid(. ~ name)
weather_plot
```
Facet grid
. = rows (nothing defines rows), don't create rows
name = define column

If reverse should have rows and not columns. 
in geompoint want to make the individual points more visible: geompoint and geomsomothe
can use alspha level to be related to other things
if you don't map that to a specific variable in the ggplot then set a univeral one in geom point


Let's combine some elements and try a new plot.

```{r}
weather_df |> 
  ggplot(aes(x=date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5)+
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name)
```

